Prediction of 2024 US election ...

### 2024-11-04 paper/paper.qmd
added description of the Trump Model

# Our Prediction

@@

We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa. Winners for other states were determined based on historical trends and predictions from sources like @cnn. Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
```{r}
#| echo: false
#| warning: false
# Split data into training and test sets

set.seed(123)

@@

train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))

@@

```{r}
#| label: fig-currentstateofelection
#| fig-cap: Kamala Harris is Predicited to be the 47th President of the United States
#| echo: false
#| warning: false

@@

```{r}

plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)

>>>>

# Trump Voter Prediction Model

The multiple linear regression model (MLR) for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris. Likewise, the Trump dataset is also split into training and testing data. The model outputs are below.

```{r}
#| echo: false
#| warning: false
set.seed(123)
train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
```

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false

just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")

predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)

predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))

result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)

# kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")

ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(color = "blue", alpha = 0.7) +
  # geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model (All Trump Data)",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```
From @fig-model-trumpmlr it is clear that the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line). Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.

```{r}
#| label: fig-model-trumpmlr-res
#| fig-cap: "MLR Trump Model does not Appear to Overfit"
#| echo: false
#| warning: false

predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)

plot(test_data_trump$pct, colMeans(predictions2), 
     xlab = "Actual PCT Values", 
     ylab = "Predicted  PCT Values", 
     main = "Predicted PCT vs Actual PCT (Test Data)")
abline(0, 1, col = "red")  # Add a 45-degree reference line

```
Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values. This suggests that model can generalization to outside data. Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.

```{r}
#| label: fig-trumpmodelci
#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
#| echo: false
#| warning: false

plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
```
Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% confidence interval ranges from around 42.5% to 47.2%. This confidence interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.

# Election Prediction

@@

We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa. Winners for other states were determined based on historical trends and predictions from sources like @cnn. Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
```{r}
#| echo: false
#| warning: false
#| label: tbl-electionprediction
#| tbl-cap: Kamala Harris Wins Most of the Swing States
# Split data into training and test sets

@@

@@

```{r}
#| label: fig-electionpredictionmap
#| fig-cap: Kamala Harris is Predicited to be the 47th President of the United States
#| echo: false
#| warning: false

@@
